---
title: "Barcode QC Report"
params:
  expRef:
    value: x,
  inputFile:
    value: x,
  samplesheet:
    value: x,
  mincnt:
    value: x,
---

```{r setup, include=FALSE}
# call with:
# R --slave -e 'rmarkdown::render(
#  input=\"${rmd_path}\",
#  output_format=\"${outformat}\",
#  output_dir=\"$PWD\",
#  params=list(expRef=\"${opt_project}\",inputFile=\"$PWD/${opt_infile}\",mincnt=\"${mincnt}\",samplesheet=\"$PWD/${opt_samplesheet}\")
#  )'

library("readr")
library("treemap")
library("ggplot2")
library("knitr")
library("kableExtra")
library("dplyr")
library("stringr")

knitr::opts_chunk$set(echo = FALSE)

# logo for document end page
nclogo <- "NC_logo_200.png"

# Convert mincnt to an integer
mincnt <- as.integer(params$mincnt)
```

# Sequel-IIe demultiplexing report for project: **`r params$expRef`**
#### Nucleomics Core, [nucleomics@vib.be]
#### date: `r Sys.Date()`

\bigskip

\bigskip

### Mosaic plot with polygon area based on barcode counts 

The read count per barcode was obtained from the regular **lima** demultiplexing analysis and used to produce the plots and tables below.

* (*`r basename(params$inputFile)`*)

```{r load data}
# load PB Lima demux counts
ccs_summary <- suppressMessages(
  read_delim(params$inputFile,
    delim = "\t",
    escape_double = FALSE,
    trim_ws = TRUE,
    show_col_types = FALSE)
)

# load samplesheet to get sample names from
samplenames <- suppressMessages(
  read_delim(params$samplesheet,
    delim = ",",
    escape_double = FALSE,
    trim_ws = TRUE,
    show_col_types = FALSE)
)

# add combined barcode string in new column
ccs_summary$BC.pairs <- paste0(ccs_summary$IdxFirstNamed, sep="--", ccs_summary$IdxCombinedNamed, sep='')

# add user provided sample names
suppressWarnings(merged_data <- merge(ccs_summary, samplenames, by.x = "BC.pairs", by.y = "Barcode", all.x = TRUE))

```

The mosaic plot of barcode counts is shown next.


```{r plot mosaic, fig.width=8, fig.height=6}
treemap(merged_data,
        index='Bio Sample',
        vSize="Counts",
        type="index",
        title="",
        fontsize.labels=8,
        lowerbound.cex.labels=0)
```

*Note:* very low barcode counts will result into small polygons without a readable label, please refer to next table for full details.

\newpage

### Barcode count distribution across all samples (N=`r nrow(ccs_summary)`)

```{r plot distribution, fig.width=8, fig.height=6}
# plot histogram
su <- summary(merged_data$Counts)
mean <- su[[4]]
median <- su[[3]]
pc5 <- quantile(merged_data$Counts,0.05)
pc95 <- quantile(merged_data$Counts,0.95)

# plot density histogram
ggplot(merged_data, aes(x=Counts)) +
  geom_histogram(bins=100, fill="grey", color="black") +
  geom_vline(aes(xintercept=mincnt), color="red", linewidth=0.75, lty=2) +
  geom_vline(aes(xintercept=mean(Counts)), color="blue", linewidth=1) +
  geom_vline(aes(xintercept=median(Counts)), color="green", linewidth=1) +
  geom_vline(aes(xintercept=pc5), color="black", linewidth=1) +
  geom_vline(aes(xintercept=pc95), color="black", linewidth=1) +
  labs(title="", x="Barcode Read Count (log-scale)", y="Frequency") +
  scale_x_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) # Log10 scale for X axis

```

**legend:** the dashed-red line indicates the threshold count of 12000 reads considered too low for proper 16S classification, black lines indicate the **bottom 5%** and **top 5%** limits [`r as.integer(pc5)`; `r as.integer(pc95)`], the blue line indicates the **mean** count value across all barcodes (`r as.integer(mean)`), the green line refers to the **median** count value (`r as.integer(median)`).

### Full result table

\footnotesize

```{r allcounts}
full_table <- merged_data[,c("Bio Sample", "BC.pairs", "Counts", "MeanScore")]
suppressWarnings(full_table$prefix <- as.numeric(sub("_.*", "", full_table$`Bio Sample`)))
full_table <- full_table[order(full_table$prefix), ]

kable(full_table[,-ncol(full_table)], row.names = FALSE,
  longtable = T) %>%
kable_styling(latex_options = c("repeat_header"),
  repeat_header_continued = "\\textit{(Continued on Next Page...)}")
```

### Samples below threshold

Drop-OFF samples (*) with **less than `r mincnt` reads** (PacBio arbitrary minimum value for proper 16S classification) are shown in the next table.

```{r low counts}
low_output <- na.omit(full_table[full_table$Counts <= mincnt, ])
kable(low_output[,-ncol(low_output)], row.names = FALSE,
  longtable = T) %>%
kable_styling(latex_options = c("repeat_header"),
  repeat_header_continued = "\\textit{(Continued on Next Page...)}")

```

(*) N=`r nrow(low_output)`

\normalsize

\bigskip

\bigskip

\bigskip

*`r Sys.Date()`*

```{r, eval=FALSE}
# Print session information
cat("R version: ", R.version.string, "\n")
cat("Session info:\n")
sessionInfo()
```

![logo](`r nclogo`) more at **<http://www.nucleomics.be>**
