---
params: 
  expRef: 
    value: x,
  inputFile:
    value: x
---

```{r setup, include=FALSE}
# call with:
# R --slave -e 'rmarkdown::render(
#  input=\"${rmd_path}\", 
#  output_format=\"${outformat}\",
#  output_dir=\"$PWD\",
#  params=list(expRef=\"${opt_project}\",inputFile=\"$PWD/${opt_infile}\")
#  )'

library("readr")
library("treemap")
library("ggplot2")
library("knitr")
library("kableExtra")
knitr::opts_chunk$set(echo = FALSE)

# logo for document end page
nclogo <- "/opt/scripts/NC_logo_200.png"
```

# Sequel-IIe demultiplexing report for project: **`r params$expRef`**
#### Nucleomics Core, [nucleomics@vib.be]
#### date: `r Sys.Date()`

\bigskip

\bigskip

The read count per barcode was obtained from the Sequel-IIe demultiplexing analysis (*`r basename(params$inputFile)`*) and used to produce the plots and tables below.

```{r load data}
# load PB demux data
ccs_summary <- read_csv(params$inputFile, show_col_types = FALSE)

# exclude undef
barcodes <- subset(ccs_summary, `Bio Sample Name` != "No Name")
```

### Mosaic plot with polygon area based on barcode counts 

```{r plot mosaic}
treemap(barcodes,
        index="Bio Sample Name",
        vSize="Polymerase Reads",
        type="index",
        title="",
        fontsize.labels=8,
        lowerbound.cex.labels=0)
```

*Note:* very low barcode counts will result into small polygons without a readable label.

### Barcode count distribution across all samples (N=`r nrow(barcodes)`)

```{r plot distribution}
# plot distribution

su <- summary(barcodes$`Polymerase Reads`)
mean <- su[[4]]
median <- su[[3]]
pc5 <- quantile(barcodes$`Polymerase Reads`,0.05)
pc95 <- quantile(barcodes$`Polymerase Reads`,0.95)

# plot density histogram
hist(barcodes$`Polymerase Reads`, breaks=100,
     main="",
     xlab="barcode read count")
abline(v=mean(barcodes$`Polymerase Reads`),col="blue",lwd=2)
abline(v=median(barcodes$`Polymerase Reads`),col="green",lwd=2)
abline(v=pc5,col="black",lwd=1)
abline(v=pc95,col="black",lwd=1)

# bottom5%
low <- barcodes[barcodes$`Polymerase Reads`<=pc5,]
low_ordered <-low[order(low$`Polymerase Reads`),]

# top5%
high <- barcodes[barcodes$`Polymerase Reads`>=pc95,]
high_ordered <- high[order(high$`Polymerase Reads`),]
```

**legend:** black lines indicate the bottom 5% and top 5% limits, the blue line indicates the **mean** count value across all barcodes (`r as.integer(mean)`), the green line refers to the **median** count value (`r as.integer(median)`).

### 5% least represented barcodes

```{r low counts}
kable(low_ordered) %>%
kable_styling(latex_options = c("striped", "hold_position"))
```

### 5% most represented barcodes

```{r higcounts}
kable(high_ordered) %>%
kable_styling(latex_options = c("striped", "hold_position"))
```

\newpage

### Full results

\footnotesize

```{r allcounts}
kable(ccs_summary[order(ccs_summary$`Polymerase Reads`),],
  longtable = T) %>%
kable_styling(latex_options = c("repeat_header"),
  repeat_header_continued = "\\textit{(Continued on Next Page...)}")
```

\normalsize

\bigskip

\bigskip

\bigskip

*`r Sys.Date()`*

![logo](`r nclogo`) more at **<http://www.nucleomics.be>**