---
params: 
  expRef: 
    value: x,
  inputFile:
    value: x
---

```{r setup, include=FALSE}
# call with:
# R --slave -e 'rmarkdown::render(
#  input=\"${rmd_path}\", 
#  output_format=\"${outformat}\",
#  output_dir=\"$PWD\",
#  params=list(expRef=\"${opt_project}\",inputFile=\"$PWD/${opt_infile}\")
#  )'

library("readr")
library("treemap")
library("ggplot2")
library("knitr")
library("kableExtra")
knitr::opts_chunk$set(echo = FALSE)

# logo for document end page
nclogo <- "/opt/scripts/NC_logo_200.png"
```

# Sequel-IIe demultiplexing report for project: **`r params$expRef`**
#### Nucleomics Core, [nucleomics@vib.be]
#### date: `r Sys.Date()`

\bigskip

\bigskip

The read count per barcode was obtained from the Sequel-IIe demultiplexing analysis (*) and used to produce the plots and tables below.

* (*`r basename(params$inputFile)`*) 

```{r load data}
# load PB Lima demux counts
ccs_summary <- suppressMessages(
  read_delim(params$inputFile, 
    delim = "\t", 
    escape_double = FALSE, 
    trim_ws = TRUE, 
    show_col_types = FALSE)
)

# add combined barcode string in new column
ccs_summary$BC.pairs <- paste0(ccs_summary$IdxFirstNamed, sep="--", ccs_summary$IdxCombinedNamed, sep='')
```

### Mosaic plot with polygon area based on barcode counts 

```{r plot mosaic}
treemap(ccs_summary,
        index="BC.pairs",
        vSize="Counts",
        type="index",
        title="",
        fontsize.labels=8,
        lowerbound.cex.labels=0)
```

*Note:* very low barcode counts will result into small polygons without a readable label, please refer to the last table for full details.

\newpage

### Barcode count distribution across all samples (N=`r nrow(ccs_summary)`)

```{r plot distribution, fig.width=5, fig.height=4}
# plot distribution

su <- summary(ccs_summary$Counts)
mean <- su[[4]]
median <- su[[3]]
pc5 <- quantile(ccs_summary$Counts,0.05)
pc95 <- quantile(ccs_summary$Counts,0.95)

# plot density histogram
hist(ccs_summary$Counts, breaks=100,
     main="",
     xlab="barcode read count")
abline(v=mean(ccs_summary$Counts),col="blue",lwd=2)
abline(v=median(ccs_summary$Counts),col="green",lwd=2)
abline(v=pc5,col="black",lwd=1)
abline(v=pc95,col="black",lwd=1)

# bottom5%
low <- ccs_summary[ccs_summary$Counts<=pc5,]
low_ordered <-low[order(low$Counts),]

# top5%
high <- ccs_summary[ccs_summary$Counts>=pc95,]
high_ordered <- high[order(high$Counts),]
```

**legend:** black lines indicate the bottom 5% and top 5% limits, the blue line indicates the **mean** count value across all barcodes (`r as.integer(mean)`), the green line refers to the **median** count value (`r as.integer(median)`).

### 5% least represented barcodes

```{r low counts}
kable(low_ordered) %>%
kable_styling(latex_options = c("striped", "hold_position"))
```

### 5% most represented barcodes

```{r higcounts}
kable(high_ordered) %>%
kable_styling(latex_options = c("striped", "hold_position"))
```

\newpage

### Full results

\footnotesize

```{r allcounts}
kable(ccs_summary[order(ccs_summary$Counts),],
  longtable = T) %>%
kable_styling(latex_options = c("repeat_header"),
  repeat_header_continued = "\\textit{(Continued on Next Page...)}")
```

\normalsize

\bigskip

\bigskip

\bigskip

*`r Sys.Date()`*

![logo](`r nclogo`) more at **<http://www.nucleomics.be>**